# Peer Assessment 1

## Loading and preprocessing the data
```{r setoption, warning=FALSE, message=FALSE}
```
### 1. Load the data
```{r cache=TRUE}
library(data.table)
data <- fread("activity.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = "NA")
data$date <- as.Date(data$date, format="%Y-%m-%d")
data$interval <- as.numeric(data$interval)

```

### 2. Remove NAs value for steps and transform date column to date value 
```{r}
data_no_mising_values <- data[!is.na(data$steps),]

```


# What is mean total number of steps taken per day?

### 1. Make a histogram of the total number of steps taken each day

```{r}
library(plyr)
total_num_steps_per_day <- ddply(data_no_mising_values, ~date, summarise, sum=sum(steps))
hist(total_num_steps_per_day$sum, main = 'the total number of steps taken each day', xlab='number of steps')
```

### 2. Calculate and report the **mean** and **median** total number of steps taken per day


```{r showtable, results='asis'}
library(plyr)
report <- ddply(data_no_mising_values,~date,summarise, mean_num_steps=mean(steps), median_num_steps=median(steps))
library(xtable)
xt <- xtable(report)
print(xt, type ="html")
```

## What is the average daily activity pattern?

### 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
mean_num_steps_per_interval <- ddply(data_no_mising_values,~interval,summarise, mean_steps=mean(steps))
plot(mean_num_steps_per_interval$interval, mean_num_steps_per_interval$mean_steps, ylab = "average number of steps taken", xlab = "Interval", type = "l")
```

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_step <- max(mean_num_steps_per_interval$mean_steps)
theInterval <- mean_num_steps_per_interval[mean_num_steps_per_interval$mean_steps == max_step, 1]
```
The 5-minute interval is `r theInterval`


## Imputing missing values

### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
total_number_of_missing_values <- nrow(data) - nrow(data_no_mising_values)
```
Number of missing values is `r total_number_of_missing_values`

### 2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Using mean_num_steps_per_interval to get the average number of steps to replace NA data 


### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
data_fill_in_missing_values <- data
data_fill_in_missing_values$steps <- mapply(function(steps, interval) 
                                                    if (is.na(steps)) 
                                                        mean_num_steps_per_interval[mean_num_steps_per_interval$interval == interval, 2]
                                                    else
                                                        steps, data_fill_in_missing_values$steps, data_fill_in_missing_values$interval)
```

### 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r , results='asis'}
library(plyr)

total_num_steps_per_day <- ddply(data_fill_in_missing_values,~date,summarise, sum=sum(steps))
hist(total_num_steps_per_day$sum, main = 'the total number of steps taken each day', xlab='number of steps')
report <- ddply(data_fill_in_missing_values,~date,summarise, mean_num_steps=mean(steps), median_num_steps=median(steps))
library(xtable)
xt <- xtable(report)
print(xt, type ="html")

```

## Are there differences in activity patterns between weekdays and weekends?

### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
data_fill_in_missing_values$wday <- as.factor(ifelse(weekdays( data_fill_in_missing_values$date) %in% c("Saturday","Sunday"), "Weekend", "Weekday")) 
```
### 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}

library(lattice)
total_num_steps_per_wday <- ddply(data_fill_in_missing_values,.(interval, wday),summarise, mean_steps=mean(steps))
xyplot(mean_steps ~ interval |  
          wday, data = total_num_steps_per_wday, 
       layout = c(1, 2), ylab = "number of steps", xlab = "Interval", type = 'l')

```

